### 2017-03-06 201703-aha-poster-2/aha-2017-cfa-slides.Rnw
initial commit programs

% ser-2017-aha-slcs.Rnw

%---------------------------------------------------------------
% Preamble
% --------------------------------------------------------------
%
% NOTE: See rice-sample.tex written by Daina Chiba at Rice University for formatting and preamble code that I copied, http://ricebeamer.dynaman.net/
\documentclass[pdf]{beamer}

\usetheme{Frankfurt}
\usecolortheme{beetle}
\setbeamersize{text margin left=10pt,text margin right=10pt} %set margin sizes

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{bm}
\usepackage{blindtext}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{csquotes}
\usepackage{booktabs}
\usepackage{caption}

\usepackage{graphicx}% http://ctan.org/pkg/graphicx

\setbeamercolor{bibliography entry title}{fg=black,bg=black}% see http://tex.stackexchange.com/questions/71352/beamer-undefined-color-local-structure
\setbeamercolor{bibliography entry location}{fg=white}
\setbeamercolor{bibliography entry note}{fg=white}
\setbeamertemplate{caption}[numbered]

% got from http://tex.stackexchange.com/questions/48023/mimic-bibtex-apalike-with-biblatex-biblatex-apa-broken
\PassOptionsToPackage{
        style=numeric,
        hyperref=true,
        backend=bibtex,
        maxbibnames=1,
        firstinits=true,
        uniquename=init,
        maxcitenames=2,
        parentracker=true,
        url=false,
        doi=true,
        isbn=false,
        eprint=false,
        backref=false,
            }{biblatex}
% see the following link for info on biblatex sort order issue: 
% http://tex.stackexchange.com/questions/51434/biblatex-citation-order
\usepackage[natbib=true, sorting=none, style=numeric, backend=biber]{biblatex}
\addbibresource{bib1}
\renewcommand*{\bibfont}{\scriptsize}

% see http://tex.stackexchange.com/questions/43083/author-year-abbr-journal-name-as-citation-style
\DeclareCiteCommand{\longcite}{(}{%
    \printnames[author]{author}, \printfield{year})}{}{}%

\newcommand{\customcite}[1]{\citeauthor{#1}, \citetitle{#1}, \citeyear{#1}}
\newcommand{\customcitetwo}[1]{\citeauthor{#1}, \citeyear{#1}}

\renewcommand{\footnotesize}{\tiny}  % change font size of citation
\renewcommand\multicitedelim{\addsemicolon\space} % This doesn't seem to work, but see http://tex.stackexchange.com/questions/167665/multiple-references-with-footfullcite

%\usepackage{fontspec} % have to compile with XeLaTeX
%\setmainfont{Arial}
\usepackage[T1]{fontenc}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault} % get something like Arial

\usepackage{amsmath,amsthm, amssymb, latexsym}
\usepackage{wrapfig}

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns

\usepackage[shortlabels]{enumitem}
% \setlist[description]{style=nextline}
\setlist[itemize]{leftmargin=0.3in, labelindent=16pt, label=$\Diamond$}

% set bullet points in itemize
% see http://tex.stackexchange.com/questions/184080/using-beamer-bullets-in-poster
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,decorations.markings,snakes,shapes}
\tikzset{
font={\fontsize{14pt}{16}\selectfont}}

\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{adjustbox}

\usepackage[framemethod=TikZ]{mdframed}
\mdfdefinestyle{MyFrame}{%
    linecolor=carolinablue,
    outerlinewidth=4pt,
    roundcorner=20pt,
    innerrightmargin=20pt,
    innerleftmargin=20pt,
    backgroundcolor=gray!10!white}

% set three column widths in poster
\newlength{\columnheight}
\setlength{\columnheight}{105cm}
\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\onecolwid}{0.31\paperwidth}
\setlength{\twocolwid}{0.31\paperwidth}
\setlength{\threecolwid}{0.31\paperwidth}

<<mainSettings, echo=FALSE, include=FALSE>>=
opts_chunk$set(echo=F, comment=NA, message=T, results = 'hide', cache=F)

library(knitr)
library(data.table)
library(readxl)
library(ggplot2)
library(plyr)
library("MplusAutomation")
library(ztable)
options(ztable.type="latex")
library(Hmisc)
library(htmlTable)
@

% --------------------------------------------------------------------------------------% 
% Title, author, date, etc.
% --------------------------------------------------------------------------------------% 
% see http://tex.stackexchange.com/questions/9740/how-can-i-add-vertical-space-to-a-beamercolorbox-to-make-it-align-with-another-o
\title{Confirmatory factor analysis (CFA) of multivariate lipid distributions across time and race-ethnic groups: United States, 2003-2012} 
\author[vonholle@unc.edu]{Ann Von Holle, Kari E. North, UNC, Chapel Hill, NC}
\institute{UNC, Chapel Hill}
\titlegraphic{\includegraphics[scale=0.25]{unc-logo.png}} %this is the path to your logo

% -------------------------------------------------------------------------------------%
% Contents
% -------------------------------------------------------------------------------------%

\begin{document}

\begin{frame}
\titlepage
\centering
\end{frame}

% \itemsep0.3em
\section{Introduction}
\subsection{Introduction}

\begin{frame}{Introduction}
          \begin{itemize}
          \normalsize  
          \item  High density lipoprotein cholesterol (HDL-C), low density lipoprotein cholesterol (LDL-C) and triglycerides (TG) are common measures of dyslipidemia.
              \begin{itemize}
                \normalsize \item Common analytic practice includes separate evaluations of each lipoprotein when evaluating dyslipidemia. 
                \item An alternative is to specify one latent factor, which predicts lipid values and incorporates their covariance structure.
                \end{itemize}
            \item Before using the latent factor, important to determine if underlying structure of dyslipidemia construct is similar across groups (measurement invariance). 
            \begin{itemize}
              \normalsize \item For example, do racial/ethnic groups have the same construct?
              \end{itemize}
            \end{itemize}
\end{frame}

        \section{Aims}
        \subsection{Aims}

        \begin{frame}{Aims}

          \begin{description}
          \item[Aim 1] Estimate dyslipidemia in a U.S. representative sample (NHANES) as one latent factor with three indicators (HDL-C, LDL-C, TG) via confirmatory factor analysis (CFA).
            \item[Aim 2] Test similarity of latent factor parameters (measurement invariance)  \footfullcite{van_de_schoot_checklist_2012} across age and racial/ethnic groups.
            \end{description}

        \end{frame}

        % -----------------------------------------------------------
        % 1-3
        % -----------------------------------------------------------
        \section{Methods}
        \subsection{Methods}

        % Figure showing{ CFA
        % %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Confirmatory Factor Analysis (CFA) Model}

          \tikzset{
            gla/.style={
          , line width=0.5mm, arrows={-Stealth[angle=45:12pt, scale=1, black, fill=black]}}}

          \tikzset{
            dgla/.style={
          , line width=0.5mm, shorten >=1pt, arrows={Stealth-Stealth[angle=45:10pt, scale=1, black, fill=black]}}}

          \tikzset{
    ultra thin/.style= {line width=0.1pt},
    very thin/.style=  {line width=0.2pt},
    thin/.style=       {line width=0.4pt},% thin is the default
    rec/.style=        {shape=rectangle, line width=0.5mm, draw},% thin is the default
    semithick/.style=  {line width=0.6pt},
    thick/.style=      {line width=0.8pt},
    very thick/.style= {line width=1.2pt},
    ultra thick/.style={line width=1.6pt}
}

\centering
        \begin{figure}
          	\begin{tikzpicture}
          		\node[line width=0.5mm, draw, ellipse, minimum height=1cm] (dys) {\large Dyslipidemia};

          		\node[above left=of dys, rec] (hdl) {\normalsize HDL-C};
          		\node[above=of dys, rec] (ldl) {\normalsize LDL-C};
          		\node[above right=of dys, rec] (tg) {\normalsize TG};

          		\node[above=of hdl, draw=none, fill=none] (e1) {\normalsize $\epsilon_1$};
          		\node[above=of ldl, draw=none, fill=none] (e2) {\normalsize $\epsilon_2$};
          		\node[above=of tg, draw=none, fill=none] (e3) {\normalsize $\epsilon_3$};

          		\draw [->, gla] (e1) to (hdl);
          		\draw [->, gla] (e2) to (ldl);
          		\draw [->, gla] (e3) to (tg);

          		\draw [->, gla] (dys) to node [anchor=east] {\large$\lambda_1$} (hdl)  ;
          		\draw [->, gla] (dys) to node [anchor=west] {\large$\lambda_2$} (ldl);
          		\draw [->, gla] (dys) to node [anchor=west] {\large$\lambda_3$} (tg);
          		\draw [<->, dgla, bend left=60]  (e1) to node [anchor=north] {\normalsize$\epsilon_{12}$} (e2);
          	\end{tikzpicture}
      \end{figure}
\end{frame}

\begin{frame}{Equations for CFA\footfullcite{brown_confirmatory_2015}}

\centering
% Equations here
\begin{array}{rcl} 
  \textrm{HDL-C} & = & \mu_1 + \lambda_1 \eta_1 + \epsilon_1 \\ 
  \textrm{LDL-C} & = & \mu_2 + \lambda_2 \eta_1 + \epsilon_2 \\ 
  \textrm{TG-C} & = & \mu_3 + \lambda_3 \eta_1 + \epsilon_3 \\ 
\end{array}

\vskip1em

Parameters:
\begin{mdframed}
\begin{description}
  \item[Factor loadings] $\lambda$, represent the association between lipids and the latent factor for dyslipidemia, $\eta_1$.
  \item[Intercepts] $\mu$, represent the means of the indicators, lipids.
  \end{description}
\end{mdframed}

\end{frame}

\begin{frame}{Four models$^a$}

\begin{itemize}
%see http://latex.wikia.com/wiki/List_of_LaTeX_symbols
  \item[\Rightarrow]\normalsize Evaluate model fit across year and racial/ethnic groups in four models, holding different parameters constant.
  \end{itemize}

\centering

  \begin{description}[align=left]
  \normalsize
    \item [Model 1]\raggedright Test metric invariance: equal factor loadings but different intercepts across groups.
    \item [Model 2]\raggedright Test intercept-only invariance: different factor loadings but equal intercepts across groups
    \item [Model 3]\raggedright Test scalar invariance: equal factor loadings and equal intercepts across groups
    \item [Model 4]\raggedright Full uniqueness invariance: equal loadings, intercept and variance across all groups
  \end{description}

\indent \raggedright \scriptsize $^a$  All models include body mass index (BMI) and age covarying with factor (not shown in figure).
      \end{frame}

        % -----------------------------------------------------------
        % 2-2
        % -----------------------------------------------------------
      \section{Results, 12-19 years}
      \subsection{Results, 12-19 years}

        % Read in code from mplus results web page development

        <<code1, cache=F>>=
          read_chunk("../mplus_results.R")
        @

        <<read1, eval=F>>=
          #NOTE: only run this as needed (with new data). Takes long time to run.
        @

        <<data-handle1>>=
        @

        <<data-handle2>>=
        @

        <<fit1>>=
        @

% see http://tex.stackexchange.com/questions/94078/resize-table-and-caption
\begin{frame}{Model fit information by groups}
\centering

\begin{minipage}[h]{1\textwidth}  
        \begin{adjustbox}{width=\linewidth} 
<<printz-1, results='asis', eval=T>>=
        print(z,type="latex")
        @
        \end{adjustbox}
\end{minipage}

\end{frame}

\begin{frame}{Best fit selections for racial/ethnic groups}

\begin{itemize} 
  \item Model 1, allowing intercepts to vary, offers best fit.
    \begin{itemize} \normalsize
      \item Acceptable levels include Root Mean Square Error of Approximation (RMSEA) < 0.05, Comparative Fit Index (CFI) = 0.95 and lowest Bayesian Information Criterion (BIC) of all models.
      \end{itemize}
    \item Evidence for metric noninvariance  ($\chi^2$(df)=155 (8), p-value<0.001) $\Rightarrow$ Cannot make meaningful comparisons across racially and ethnically diverse populations.

  \end{itemize}
\end{frame}

\begin{frame}{Best fit selections for time groups}

  \begin{itemize}
    \item Model 1, allowing intercepts to vary, offers best fit.
    \begin{itemize} \normalsize
      \item RMSEA < 0.05 and lowest Bayesian Information Criterion (BIC) of all models.
      \end{itemize}
    \item Evidence for metric noninvariance ($\chi^2$(df)=155 (12), p-value<0.001)  $\Rightarrow$ Cannot make meaningful comparisons across time.
    \end{itemize}
\end{frame}

% Second table here
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<loadings1>>=
@

<<combo1>>=
@

<<print-combo-race,  warning=TRUE>>=
@

\begin{frame}{Model 1 standardized parameters by race/ethnic groups}

\centering
\begin{minipage}[h]{0.9\textwidth}                          % minipage spans half the textwidth
        \begin{adjustbox}{width=\linewidth}  % can also use \linewidth or sth. else
<<printz-3, results='asis', eval=T>>=
        print(z,type="latex")
        @
        \end{adjustbox}
\end{minipage}
\end{frame}

\begin{frame}{Latent factor characteristics by racial/ethnic groups}

          \begin{itemize}
            \item \underline{Factor loadings}: A one unit factor increase represents less favorable lipid value $\Rightarrow$ an increase in LDL-C (0.4) and TG (0.6) accompanied by a decrease in HDL ($\sim$ -0.6).
            \item \underline{Intercepts}: Non-Hispanic Black group only one with positive HDL-C (0.132) and lowest TG (-1.078) relative to center of multivariate distribution.
          \end{itemize}
\end{frame}

<<print-combo-year,  warning=TRUE>>=
@

\begin{frame}{Model 1 standardized parameters by year groups}

\centering
\begin{minipage}[h]{0.8\textwidth}                          % minipage spans half the textwidth
        \begin{adjustbox}{width=\linewidth}  % can also use \linewidth or sth. else
<<printz-4, results='asis', eval=T>>=
        print(z,type="latex")
        @
        \end{adjustbox}
\end{minipage}
\end{frame}

% NOTE: not enough space here for gender. Remove to allow more
% discussion of results in poster. keep for slides

\begin{frame}{Fit and parameters for groups by sex}

\vskip-1em

% Run code for gender groups (from mplus_results.R)
<<fit-gender>>=
@
\centering
\begin{adjustbox}{width=0.7\linewidth} 
%\caption*{Model fit}
<<printf-gender, results='asis', eval=T>>=
        print(z,type="latex",
              caption="Model fit")
@
\end{adjustbox}

\vskip1em

<<print-combo-gender>>=
@
\centering
\begin{adjustbox}{width=0.7\linewidth}
%\caption*{Model 1 parameters}
<<printp-gender, results='asis', eval=T>>=
        print(z,type="latex",
              caption="Model 1 parameters")
@
\end{adjustbox}

\vskip1em

\raggedright \footnotesize Invariance testing indicates evidence for metric non-invariance when testing difference between a metric model (model 1) and a full invariance model (model 4) in which there are full equality constraints across groups ($\chi^2$ = 68.5 (df=2); p-value<0.001).

% NOTE: below is column approach, but too small to read

% \begin{columns}[T] % align columns
% \begin{column}{.48\textwidth}
% 
% %Left Part -------------------
% 
% % Run code for gender groups (from mplus_results.R)
% <<fit-gender>>=
% @
% \centering
% \begin{minipage}[h]{1\textwidth}  
% \begin{adjustbox}{width=\linewidth} 
% %\caption*{Model fit}
% <<printf-gender, results='asis', eval=T>>=
%         print(z,type="latex",
%               caption="Model fit")
% @
% \end{adjustbox}
% \end{minipage}
% 
% 
% \end{column}%
% 
% \hfill%
% \begin{column}{.48\textwidth}
% 
% 
% % Right Part --------------------
% 
% <<print-combo-gender>>=
% @
% \centering
% \begin{minipage}[h]{1\textwidth}  
% \begin{adjustbox}{width=\linewidth}
% %\caption*{Model 1 parameters}
% <<printp-gender, results='asis', eval=T>>=
%         print(z,type="latex",
%               caption="Model 1 parameters")
% @
% \end{adjustbox}
% \end{minipage}
% 
% \end{column}%
% \end{columns}

\end{frame}

% >>>>>>>>>>>>>>>>>>>>>>>>>>>

\begin{frame}{Model fitting process}
          \begin{itemize}
          \item Covariance between HDL-C and TG no longer significant when accounting for BMI covarying with latent lipid factor.
          \item Age positive covariance with lipid factor.
          \item BMI positive covariance with lipid factor.
          \end{itemize}
\end{frame}

\section{Summary}
\subsection{Summary}

\begin{frame}{Summary} 
\Large
            \begin{itemize}
                \item 3-indicator latent factor can function as a parsimonious and well-fitting measure for dyslipidemia in structural equation models.
                \item []
                \item Important not to compare latent factor characteristics across groups by race/ethnic, time or sex status given metric measurement non-invariance.
            \end{itemize}
\end{frame}

\begin{frame}

Contact information: vonholle at unc dot edu

Further project information: https://avonholle.github.io/cfa-lipids/

\end{frame}

\end{document}