### 2017-03-06 201703-aha-poster-2/aha-2017-cfa.Rnw

% aha-2017-slcs.Rnw

%---------------------------------------------------------------
% Preamble
% --------------------------------------------------------------
%
% NOTE: See rice-sample.tex written by Daina Chiba at Rice University for formatting and preamble code that I copied, http://ricebeamer.dynaman.net/
\documentclass[final]{beamer}
\usepackage[orientation=landscape, size=custom, width=121.92, height=106.68, scale=1.7]{beamerposter}  % this matches 4 feet (48 inches) by 3.5 feet (42 inches). Will get from phdposters.com 1 inch per 2.54 cm

\mode<presentation>{\usetheme{UNC5}}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{bm}
\usepackage{blindtext}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{csquotes}

\setbeamercolor{bibliography entry title}{fg=black,bg=black}% see http://tex.stackexchange.com/questions/71352/beamer-undefined-color-local-structure
\setbeamertemplate{caption}[numbered]

% got from http://tex.stackexchange.com/questions/48023/mimic-bibtex-apalike-with-biblatex-biblatex-apa-broken
\PassOptionsToPackage{
        style=numeric,
        hyperref=true,
        backend=bibtex,
        maxbibnames=1,
        firstinits=true,
        uniquename=init,
        maxcitenames=2,
        parentracker=true,
        url=false,
        doi=true,
        isbn=false,
        eprint=false,
        backref=false,
            }{biblatex}
% see the following link for info on biblatex sort order issue: 
% http://tex.stackexchange.com/questions/51434/biblatex-citation-order
\usepackage[natbib=true, sorting=none, style=numeric, backend=biber]{biblatex}
\addbibresource{bib1}
\renewcommand*{\bibfont}{\scriptsize}

%\usepackage{fontspec} % have to compile with XeLaTeX
%\setmainfont{Arial}
\usepackage[T1]{fontenc}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault} % get something like Arial

\usepackage{amsmath,amsthm, amssymb, latexsym}
\usepackage{wrapfig}

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns

\usepackage[shortlabels]{enumitem}
% \setlist[description]{style=nextline}
\setlist[itemize]{leftmargin=0.6in, labelindent=16pt, label=$\Diamond$}

% set bullet points in itemize
% see http://tex.stackexchange.com/questions/184080/using-beamer-bullets-in-poster
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,decorations.markings,snakes,shapes}
\tikzset{
font={\fontsize{14pt}{16}\selectfont}}

\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{adjustbox}

\usepackage[framemethod=TikZ]{mdframed}
\mdfdefinestyle{MyFrame}{%
    linecolor=carolinablue,
    outerlinewidth=4pt,
    roundcorner=20pt,
    innerrightmargin=20pt,
    innerleftmargin=20pt,
    backgroundcolor=gray!10!white}

% set three column widths in poster
\newlength{\columnheight}
\setlength{\columnheight}{105cm}
\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\onecolwid}{0.31\paperwidth}
\setlength{\twocolwid}{0.31\paperwidth}
\setlength{\threecolwid}{0.31\paperwidth}

<<mainSettings, echo=FALSE, include=FALSE>>=
opts_chunk$set(echo=F, comment=NA, message=T, results = 'hide', cache=F)

library(knitr)
library(data.table)
library(readxl)
library(ggplot2)
library(plyr)
library("MplusAutomation")
library(ztable)
options(ztable.type="latex")
library(Hmisc)
library(htmlTable)
@

% --------------------------------------------------------------------------------------% 
% Title, author, date, etc.
% --------------------------------------------------------------------------------------% 
% see http://tex.stackexchange.com/questions/9740/how-can-i-add-vertical-space-to-a-beamercolorbox-to-make-it-align-with-another-o
\title{Confirmatory factor analysis (CFA) of multivariate lipid distributions across time and race-ethnic groups: United States, 2003-2012} 
\author[vonholle@unc.edu]{Ann Von Holle, Kari E. North, UNC, Chapel Hill, NC}
\institute{UNC}
\titlegraphic{unc-black.eps} %this is the path to your logo

% -------------------------------------------------------------------------------------%
% Contents
% -------------------------------------------------------------------------------------%

\begin{document}

\begin{frame}[t]

  \begin{columns}[T] % t instead of T or c means columns start at top

    % ---------------------------------------------------------%
    % Set up 1st column
    % ---------------------------------------------------------%
    \begin{column}{\onecolwid}
    \begin{beamercolorbox}[wd=\textwidth]{postercolumn}
    % fill each column with content
        % -----------------------------------------------------------
        % 1-1 (first column's first block
        % -----------------------------------------------------------
        % fill each column with content
        \begin{block}{Introduction}

          \begin{itemize}
%          \itemsep0.3em
          \normalsize  
          \item  High density lipoprotein cholesterol (HDL-C), low density lipoprotein cholesterol (LDL-C) and triglycerides (TG) are common measures of dyslipidemia.
              \begin{itemize}
                \normalsize \item Common analytic practice includes separate evaluations of each lipoprotein when evaluating dyslipidemia. 
                \item An alternative is to specify one latent factor, which predicts lipid values and incorporates their covariance structure.
                \end{itemize}
            \item Before using the latent factor, important to determine if underlying structure of dyslipidemia construct is similar across groups (measurement invariance). 
            \begin{itemize}
              \normalsize \item For example, do racial/ethnic groups have the same construct?
              \end{itemize}
            \end{itemize}

        \end{block}
        \vskip1ex

        % -----------------------------------------------------------
        % 1-2
        % -----------------------------------------------------------
        \begin{block}{Aims}

          \begin{enumerate}[1.]
          \item Estimate dyslipidemia in a U.S. representative sample (NHANES) as one latent factor with three indicators (HDL-C, LDL-C, TG) via confirmatory factor analysis (CFA).
            % \begin{itemize}
            %   \item Compare model fit using common indicators.
            %   \end{itemize}
            \item Test similarity of latent factor parameters (measurement invariance) across age and racial/ethnic groups.
            % \begin{itemize}
            %   \normalfont \item Groups include racial/ethnic status and time (years).
            %   \end{itemize}
            \end{enumerate}

        \end{block}
        \vskip1ex

        % -----------------------------------------------------------
        % 1-3
        % -----------------------------------------------------------
        \begin{block}{Methods: CFA}

        % Figure showing CFA
        % %%%%%%%%%%%%%%%%%%%%%%%%%%%%

          \tikzset{
            gla/.style={
          , line width=1mm, arrows={-Stealth[angle=45:12pt, scale=2, black, fill=black]}}}

          \tikzset{
            dgla/.style={
          , line width=1mm, shorten >=1pt, arrows={Stealth-Stealth[angle=45:10pt, scale=1.5, black, fill=black]}}}

          \tikzset{
    ultra thin/.style= {line width=0.1pt},
    very thin/.style=  {line width=0.2pt},
    thin/.style=       {line width=0.4pt},% thin is the default
    rec/.style=        {shape=rectangle, line width=1mm, draw},% thin is the default
    semithick/.style=  {line width=0.6pt},
    thick/.style=      {line width=0.8pt},
    very thick/.style= {line width=1.2pt},
    ultra thick/.style={line width=1.6pt}
}

\vskip-1ex
\centering
        \begin{figure}
           \scalebox{1.2}{
          	\begin{tikzpicture}

          		\node[line width=1mm, draw, ellipse, minimum height=1cm] (dys) {\large Dyslipidemia};

          		\node[above left=of dys, xshift=2cm, yshift=4cm, rec] (hdl) {\normalsize HDL-C};
          		\node[above=of dys, yshift=4cm, rec] (ldl) {\normalsize LDL-C};
          		\node[above right=of dys, xshift=-2cm, yshift=4cm, rec] (tg) {\normalsize TG};

          		\node[above=of hdl, draw=none, fill=none] (e1) {\normalsize $\epsilon_1$};
          		\node[above=of ldl, draw=none, fill=none] (e2) {\normalsize $\epsilon_2$};
          		\node[above=of tg, draw=none, fill=none] (e3) {\normalsize $\epsilon_3$};

          		\draw [->, gla] (e1) to (hdl);
          		\draw [->, gla] (e2) to (ldl);
          		\draw [->, gla] (e3) to (tg);

          		\draw [->, gla] (dys) to node [anchor=east] {\large$\lambda_1$} (hdl)  ;
          		\draw [->, gla] (dys) to node [anchor=west] {\large$\lambda_2$} (ldl);
          		\draw [->, gla] (dys) to node [anchor=west] {\large$\lambda_3$} (tg);
          		\draw [<->, dgla, bend left=60]  (e1) to node [anchor=north] {\normalsize$\epsilon_{12}$} (e2);

          	\end{tikzpicture}
           }
      \end{figure}

Equations for CFA:\left\{
% Equations here
\begin{array}{rcl} 
  \textrm{HDL-C} & = & \mu_1 + \lambda_1 \eta_1 + \epsilon_1 \\ 
  \textrm{LDL-C} & = & \mu_2 + \lambda_2 \eta_1 + \epsilon_2 \\ 
  \textrm{TG-C} & = & \mu_3 + \lambda_3 \eta_1 + \epsilon_3 \\ 
\end{array}

\begin{itemize}
%see http://latex.wikia.com/wiki/List_of_LaTeX_symbols
  \item[\Rightarrow]\normalsize Evaluate model fit across year and racial/ethnic groups in four models, holding different parameters constant.
  \end{itemize}

        \end{block}

  \end{beamercolorbox}
  \end{column}
    % ---------------------------------------------------------%
    % end the 1st column
    % ---------------------------------------------------------%

% ---------------------------------------------------------%
% Set up 2nd column
% ---------------------------------------------------------%

\begin{column}{\twocolwid}
\begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}

        % -----------------------------------------------------------
        % 2-1
        % -----------------------------------------------------------
      \begin{block}{Methods, cont...}

\begin{itemize}
  \item Factor loadings, $\lambda$, represent the association between lipids and the latent factor for dyslipidemia, $\eta_1$.
  \item Intercepts, $\mu$, represent the means of the indicators, lipids.
  \end{itemize}

\centering
\textbf{Four models$^a$}
\begin{mdframed}[style=MyFrame]
  \begin{description}[align=left]
  \normalsize
    \item [Model 1]\raggedright Test metric invariance: equal factor loadings but different intercepts across groups.
    \item [Model 2]\raggedright Test intercept-only invariance: different factor loadings but equal intercepts across groups
    \item [Model 3]\raggedright Test scalar invariance: equal factor loadings and equal intercepts across groups
    \item [Model 4]\raggedright Full uniqueness invariance: equal loadings, intercept and variance across all groups
    \end{description}
\end{mdframed}
\indent \raggedright \scriptsize $^a$  All models include body mass index (BMI) and age covarying with factor (not shown in figure).
      \end{block}
      \vskip1ex

        % -----------------------------------------------------------
        % 2-2
        % -----------------------------------------------------------
      \begin{block}{Results, 12-19 years}

        % Read in code from mplus results web page development

        <<code1, cache=F>>=
          read_chunk("../mplus_results.R")
        @

        <<read1, eval=F>>=
          #NOTE: only run this as needed (with new data). Takes long time to run.
        @

        <<data-handle1>>=
        @

        <<data-handle2>>=
        @

        <<fit1>>=
        @

% see http://tex.stackexchange.com/questions/94078/resize-table-and-caption

\centering
Model fit information

\begin{mdframed}[style=MyFrame]
\begin{minipage}[h]{1\textwidth}  % minipage spans half the textwidth
        \begin{adjustbox}{width=\linewidth}  % can also use \linewidth or sth. else
<<printz-1, results='asis', eval=T>>=
        print(z,type="latex")
        @
        \end{adjustbox}
\end{minipage}
\end{mdframed}

\textbf{Racial/ethnic groups}
\begin{itemize} 
  \item Model 1, allowing intercepts to vary, offers best fit.
    \begin{itemize} \normalsize
      \item Acceptable levels include Root Mean Square Error of Approximation (RMSEA) < 0.05, Comparative Fit Index (CFI) = 0.95 and lowest Bayesian Information Criterion (BIC) of all models.
%  \item No evidence model 1 different than configural model relaxing all parameter assumptions ($\chi^2$=9.12, p-value=0.33, df=8).
      \end{itemize}
  \end{itemize}

\textbf{Time groups}
  \begin{itemize}
    \item Model 1, allowing intercepts to vary, offers best fit.
    \begin{itemize} \normalsize
      \item RMSEA < 0.05, CFI = 0.93 and lowest Bayesian Information Criterion (BIC) of all models.
      %\item Evidence suggests metric and scalar invariance.% (all $\chi^2$ tests against configural model not significant).
      \end{itemize}
    \end{itemize}

% Second table here
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<loadings1>>=
@

<<combo1>>=
@

\end{block}

\end{beamercolorbox}
\end{column}

% ---------------------------------------------------------%
% end the 2nd column

% ---------------------------------------------------------%
% Set up 3rd column
% ---------------------------------------------------------%

\begin{column}{\threecolwid}
\begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}

        \vfill

        % -----------------------------------------------------------
        % 3-1
        % -----------------------------------------------------------
%NOTE: results here need to be finessed in Mplus. will work on later.

%         \begin{block}{Results, 20-50 years}
% 
% 
% <<fit2>>=
% @
%  
% % see http://tex.stackexchange.com/questions/94078/resize-table-and-caption
% \centering
% Model fit information
% \vskip0.5ex
% 
% \begin{mdframed}[style=MyFrame]
% \begin{minipage}[h]{1\textwidth}                          % minipage spans half the textwidth
%         \begin{adjustbox}{width=\linewidth}  % can also use \linewidth or sth. else
% <<printz-4, results='asis', eval=T>>=
%         print(z,type="latex")
%         @
%         \end{adjustbox}
% \end{minipage}
% \end{mdframed}
% \vskip1ex
% 
%         <<print-combo-race-50>>=
%         @
% 
% \centering
% Model 1 parameters for race/ethnic groups
% 
% \begin{mdframed}[style=MyFrame]
% \begin{minipage}[h]{1.0\textwidth}                          % minipage spans half the textwidth
%         \begin{adjustbox}{width=\linewidth}  % can also use \linewidth or sth. else
% <<printz-5, results='asis', eval=T>>=
%         print(z,type="latex")
%         @
%         \end{adjustbox}
% \end{minipage}
% \end{mdframed}
% 
%         \end{block}
% 
%         \vskip1ex
%         \vfill

        % -----------------------------------------------------------
        % 3-2
        % -----------------------------------------------------------
\begin{block}{Results, 12-19 years, cont...}

\centering\normalsize\textbf{Lipid construct comparisons across:}

%\centering\normalsize \textbf{Across racial/ethnic groups} %see https://en.wikibooks.org/wiki/LaTeX/Colors
  \begin{description}\normalsize
                \item[Racial/ethnic groups] Evidence for metric non-invariance ($\chi^2$(df)=155 (8), p-value<0.001) $\Rightarrow$ Cannot make meaningful comparisons across racially and ethnically diverse populations.
                \item[Time (years)] Evidence for metric non-invariance  ($\chi^2$(df)=155 (12), p-value<0.001)  $\Rightarrow$ Cannot make meaningful comparisons across time.
      \end{description}
\vskip0.5em
% See examine-loadings-rev-chi-sq-tests.xlsx for Satorra-Bentler Scaled Chi-Square test (https://www.statmodel.com/chidiff.shtml)

<<print-combo-race,  warning=TRUE>>=
@

\centering
\normalsize Model 1 standardized parameters by race/ethnic groups

\begin{mdframed}[style=MyFrame]
\begin{minipage}[h]{1.0\textwidth}                          % minipage spans half the textwidth
        \begin{adjustbox}{width=\linewidth}  % can also use \linewidth or sth. else
<<printz-3, results='asis', eval=T>>=
        print(z,type="latex")
        @
        \end{adjustbox}
\end{minipage}
\end{mdframed}

\normalsize\centering\textbf{Latent factor characteristics:}
  \begin{description}\normalsize
            \item[Factor loadings] A one unit factor increase represents less favorable lipid value $\Rightarrow$ an increase in LDL-C (0.4) and TG (0.6) accompanied by a decrease in HDL ($\sim$ -0.6).
            \item[Intercepts] Non-Hispanic Black group only one with positive HDL-C (0.132) and lowest TG (-1.078) relative to center of multivariate distribution.
          \end{description}

% Model fitting process
%           \begin{itemize}
%           \item Covariance between HDL-C and TG no longer significant when accounting for BMI covarying with latent lipid factor.
%           \item Age positive covariance with lipid factor.
%           \item BMI positive covariance with lipid factor.
%           \end{itemize}

        \end{block}
        \vskip1ex

        % -----------------------------------------------------------
        % 3-3
        % -----------------------------------------------------------
        \begin{block}{Summary}

            \begin{itemize}
                \item[\Rightarrow] 3-indicator latent factor can function as a parsimonious and well-fitting measure for dyslipidemia in structural equation models.
                \item[\Rightarrow] Cannot compare latent factor characteristics across groups by race/ethnic or time status.
            \end{itemize}

        \end{block}
        \vskip1ex
        \vfill

        % -----------------------------------------------------------
        % 3-3
        % -----------------------------------------------------------
        \begin{block}{References}
          %\scriptsize{%          \printbibliography{}}
          \scriptsize
          (1) \fullcite{van_de_schoot_checklist_2012} \\
          (2) \fullcite{brown_confirmatory_2015}
          % see https://en.wikibooks.org/wiki/LaTeX/Fonts
        \end{block}

\end{beamercolorbox}
\end{column}
% ---------------------------------------------------------%
% end the 3rd column
% ---------------------------------------------------------%

\end{columns}

\end{frame}
\end{document}