### 2017-03-06 mplus-check1.Rmd

---
title: "Mplus: sensitivity analysis for useobservations vs subpopulation command"
author: "Ann Von Holle"
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: lit.bib
output:
  html_document:
        toc: yes
        toc_float: yes
        number_sections: true
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(knitr)
opts_chunk$set(fig.width=10, fig.height=10, 
               fig.keep = 'all',
               fig.align="left", 
               echo=F, results='hide',
               warning=FALSE, message=FALSE, comment=NA)

library("MplusAutomation")
library(data.table)
```

<!-- Run mplusautomate package to extract results from mplus output, .out 
 See https://cran.r-project.org/web/packages/MplusAutomation/vignettes/Vignette.pdf
-->

```{r}
# run analyses with subpopulation (no grouping)

runModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/sensitivity', replaceOutfile="never")

```

```{r}
# read in parameter estimates from model using subpopulation and useobservatios options.

# extract out all parameters from output
sub.params <- extractModelParameters("C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/sensitivity", recursive=TRUE)

names(sub.params)
names(sub.params) = c("sub", "use") # rename to shorter name

```

```{r}
# Compare model parameters (and se). Look for differences.

names(sub.params$sub$unstandardized)
sub = data.frame(sub.params$sub$unstandardized)
use = data.frame(sub.params$use$unstandardized)

sub

# make a new value to put in table: est (se)
sub = within(sub, {
  est.se = paste0(est, " (", se, ")")
  option="sub"
})

use = within(use, {
  est.se = paste0(est, " (", se, ")")
  option="use"
})

tot = rbind(sub, use)

pct.diff.se = 100*((use$se - sub$se)/sub$se)

t1 = data.table(dcast(setDT(tot), paramHeader + param ~ option, 
                      value.var = "est.se"))

t1$pct.diff.se = round(pct.diff.se,2)
t1

colnames(t1) = c("parameter", "lipid indicator", "est(se) from sub", 
                 "est(se) from use", "% diff se")
```

# Table comparing parameter results using SUBPOPULATION option in complex design versus USEOBSERVATIONS option.

```{r, results='markup'}
  kable(t1)
```

Problem with complex design and GROUPING option: To determine measurement invariance I need to use the grouping option. However, with you can't use the SUBPOPULATION option and GROUPING command together. So I am checking to see if there is a big difference when using useobservations instead of subpopulation. Based on findings in table there is not a big difference. If anything the standard errors appear to be slightly larger if there is any chance at all.