### 2017-03-06 mplus-results.Rmd

---
title: "Mplus Results"
author: "Ann Von Holle"
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: lit.bib
output:
  html_document:
        toc: yes
        toc_float: yes
        number_sections: true
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(knitr)
opts_chunk$set(fig.width=10, fig.height=10, 
               fig.keep = 'all',
               fig.align="left", 
               echo=F, results='hide',
               warning=FALSE, message=FALSE, comment=NA)

library("MplusAutomation")
library(data.table)
library(plyr)

library(ztable)
options(ztable.type="html")
#options(ztable.type="viewer") # for debugging ztable

```

```{r, eval=FALSE}
# generate data to read in to Mplus
  read_chunk("export-mplus.R")
```

```{r}
# Look at some descriptive statistics in data

load("alldat.Rda") # get appended data (all.dat.df) from export-mplus.R
summary(all.dat.df)

# get parameters to set in mplus file templates for variance of bmi and age

# mean, sd bmi
with(all.dat.df, {
  c(mean(bmi, na.rm=T),
    sd(bmi, na.rm=T),
    mean(age, na.rm=T),
    sd(age, na.rm=T))})

table(all.dat.df$age.cut)

# get mean bmi and age by age cut groups
all.dt = data.table(all.dat.df[c("bmi", "age", "age.cut")])
all.dt[, sapply(.SD, function(x) list(mean=mean(x, na.rm=T), 
                                      sd=sd(x, na.rm=T))), 
       by=age.cut]

```

<!-- Make the models from an MplusAutomate template -->

```{r}

# Models comparing racial/ethnic groups and time (years)

# Each template iterates across age groups and groupings (year and race)
# terms from model come from "A checklist for testing measurement invariance",
# doi: 10.1080/17405629.2012.686740
# ----------------------------------------------------------------

# Template for Model 1: metric invariance. 
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m1.txt')

# Template for Model 2: Intercept only invariance
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m2.txt')

# Template for Model 3: Scalar invariance
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m3.txt')

# Template for Model 4: Full uniqueness invariance
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m4.txt')

# NOTE: after creating the models go to virtuallab and run them in MPlus. Details below.cc
```

```{r}

# Models comparing gender groups

# Each template iterates across age groups

# terms from model come from "A checklist for testing measurement invariance",
# doi: 10.1080/17405629.2012.686740
# ----------------------------------------------------------------

# Template for Model 1: metric invariance. 
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m1_sex.txt')

# Template for Model 2: Intercept only invariance
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m2_sex.txt')

# Template for Model 3: Scalar invariance
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m3_sex.txt')

# Template for Model 4: Full uniqueness invariance
createModels('C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template/template_m4_sex.txt')

```

<!-- Run models created from template.
    Notes:
    1. I originally worked out all the problems in files for individual models/groups and age groups in folder titled, 
    ~\Dropbox\unc.grad.school\my-papers\ms-201608-2\programs-ms201608-2\mplus\archive\ 
    2. this won't run for the virtual lab Mplus. trying to figure out the command line to run mplus 
-->

```{r, eval=FALSE}

# -----------------------------------
# Note: I was able to run batches of Mplus files from the virtuallab.unc.edu server
# after emailing tech support 9/23/2016. A special version of R on virtuallab
# was set up for me to be able to call Mplus in virtuallab. was able to run
# all 28 models in less than 5 minutes after that.

# Contents of run-models.R program below (but note that it has to be run in R while in virtuallab.unc.edu). Before running run-models.R I run the chunk of code above to get the .inp files on my local folder (c:\temp\models\). then I copy those .inp files to the c:\temp\models folder. next I copy and paste the c:\temp\models\ folder on my local computer to my temporary folder in virtuallab.unc.edu server using their version of windows explorer and typing "%userprofile%". this pulls up my temporary server folder that R in virutal lab can locate. I also copy and paste the totdatlong.dat file made in 'export-mplus.R' (also in c:\temp\models\ folder) to the c:\users\vonholle\virtual-lab\ folder (or whatever you name it to in the virtual lab desktop, could be c:\users\vonholle\temp) to folder holding my programs in virtuallab.unc.edu. Then I run the run-models.R program and it should create .out files in the c:\users\vonholle\virtual-lab\ folder. Then I copy and paste those .out files back to my local drive (at ~\Dropbox\unc.grad.school\my-papers\ms-201608-2\programs-ms201608-2\mplus\compare-years-template) so I can run the rest of this .Rmd file.
# -----------------------------------

# install.packages("MplusAutomation")
# library(MplusAutomation)
# 
# runModels("C:\\Users\\vonholle\\temp")

```

<!-- Extract results from mplus output, .out 
 See https://cran.r-project.org/web/packages/MplusAutomation/vignettes/Vignette.pdf
-->

```{r read1, cache=F}

# read in results for differences across years (all ages) and racial/ethnic groups
# NOTE: I have to move the files from c:\temp\models\ to the folder specified below

# extract out all parameters from output
amp.years.params <- extractModelParameters("C:/temp/models2", recursive=F)
warnings() # not sure why there are some problems reading in data -- missing values?

#NOTE: be careful with the recursive setting. will get repeats that mess up tables below.

ayp = amp.years.params # save a backup because it takes so long to run
names.ayp = names(ayp)

# extract out model fit summaries from output
amp.years.sums <- extractModelSummaries("C:/Users/vonholle/Dropbox/unc.grad.school/my-papers/ms-201608-2/programs-ms201608-2/mplus/compare-years-template", recursive=T)
names(amp.years.sums)
```

```{r data-handle1}

# check names of objects in list
names(amp.years.params) # names too long, simplify

nms = names.ayp

gender.g = ifelse(grepl("sex.all.", nms)==T, 'all',
                  ifelse(grepl("sex.fem.", nms)==T, 'female',
                         ifelse(grepl("sex.mal.", nms)==T, 'male', 
                                ifelse(grepl("groups.gender", nms)==T, 'gen', NA))))
table(gender.g) # check

ageg = ifelse(grepl("12to19",nms)==T, "12to19",
              ifelse(grepl("20to50",nms)==T,"20to50",
                     ifelse(grepl("50to80",nms)==T, "50to80", NA)))

table(ageg) #check

gps = ifelse(grepl("groups.year",nms)==T, "YR",
             ifelse(grepl("groups.race", nms)==T, "RE",
                    ifelse(grepl("groups.gender", nms)==T, "GE", NA)))
table(gps)

model = ifelse(grepl(".m1", nms)==T, 1,
               ifelse(grepl(".m2", nms)==T, 2,
                      ifelse(grepl(".m3", nms)==T, 3,
                             ifelse(grepl(".m4", nms)==T, 4, NA))))
table(model)

names(amp.years.params) = paste0("m", model, "y", ageg, "g", gps, "gender", gender.g)
names(amp.years.params) #check
head(amp.years.params)
```

```{r data-handle2}
# --------------------------------------------
# PREP model fit info
# --------------------------------------------

head(amp.years.sums)

# add descriptors to model fit data

amp.years.sums$gender.g = with(amp.years.sums, {
  ifelse(grepl("-all-", Filename)==T, 'all',
                  ifelse(grepl("-female-", Filename)==T, 'female',
                         ifelse(grepl("-male-", Filename)==T, 'male', NA)))
})

table(amp.years.sums$gender.g) # check

amp.years.sums$agegrp = with(amp.years.sums, {
  ifelse(grepl("12to19",Filename)==T, "12to19",
                                 ifelse(grepl("20to50",Filename)==T,"20to50",
                                        ifelse(grepl("50to80",Filename)==T, "50to80", NA)))
})
table(amp.years.sums$agegrp)

amp.years.sums$group = with(amp.years.sums, {
  ifelse(grepl("groups.year",nms)==T, "YR",
             ifelse(grepl("groups.race", nms)==T, "RE",
                    ifelse(grepl("groups.gender", nms)==T, "GE", NA)))
})

table(amp.years.sums$group)
names(amp.years.sums)

amp.years.sums$model = with(amp.years.sums, {
  ifelse(grepl("m1-", Filename)==T, 1,
               ifelse(grepl("m2-", Filename)==T, 2,
                      ifelse(grepl("m3-", Filename)==T, 3,
                             ifelse(grepl("m4-", Filename)==T, 4, NA))))
})

table(amp.years.sums$model)

# Get columns to print off table below
names(amp.years.sums)
head(amp.years.sums)
fit.vals = c("Filename",
             "ChiSqM_Value", "ChiSqM_DF", "ChiSqM_PValue", 
             "LL",
             "CFI", "TLI", "RMSEA_Estimate",
             "BIC", "AIC", "gender.g", "agegrp", "group", "model")

fit.vals.renamed = c("Chi-sq", "df", "p-value", "LL",
                     "CFI", "TLI",
                     "BIC", "AIC", "RMSEA", "Model", "Gender", "Age", "Groups", "mod")

amp.years.sums.sub = amp.years.sums[,colnames(amp.years.sums) %in% fit.vals]
colnames(amp.years.sums.sub) = fit.vals.renamed

head(amp.years.sums.sub) #check

# make a model description to describe model invariance settings as described in
# van de Schoot R, Lugtig P, Hox J. A checklist for testing measurement invariance. European Journal of Developmental Psychology. 2012 Jul;9(4):486–92.

amp.years.sums.sub = within(amp.years.sums.sub, {
  description = ifelse(grepl("m1", Model)==T, "M1: Metric invariance: diff int",
                       ifelse(grepl("m2", Model)==T, "M2: Diff factor loadings",
                              ifelse(grepl("m3", Model)==T, "M3: Scalar Invariance: eq loading and eq int",
                                     ifelse(grepl("m4", Model)==T, "M4: Full Uniqueness invariance: eq loadings, intercepts and res variance",
                                            ifelse(grepl("m5", Model)==T, "M5: Free loadings, int and res variance",
                                                   NA)))))
})

table(amp.years.sums.sub$Groups)

order = amp.years.sums.sub[order(amp.years.sums.sub$Age,
                         amp.years.sums.sub$Groups,
                         amp.years.sums.sub$description),]

  mod.1 = order[order$Gender=="all" & order$Groups %in% c("RE", "YR"), !(colnames(order) %in% c("Model"))]
  table(mod.1$Groups)
  head(mod.1)

  mod.1f = order[order$Gender=="female"& order$Groups %in% c("RE", "YR"),
                 !(colnames(order) %in% c("Model"))]
  head(mod.1f)
  table(mod.1f$Groups)

  mod.1m = order[order$Gender=="male"& order$Groups %in% c("RE", "YR"),
                 !(colnames(order) %in% c("Model"))]
  head(mod.1m)

  mod.ge = order[order$Groups %in% c("GE"), !(colnames(order) %in% c("Model"))]
  head(mod.ge)

  mod.1.tot = rbind(mod.1[complete.cases(mod.1$df),], 
                    mod.ge) #includes gender groups (in addition to race/ethnic and year groups)

  table(mod.1.tot$Groups)
  mod.1.tot = mod.1.tot[order(mod.1.tot$Groups, mod.1.tot$Age, mod.1.tot$description),]
  mod.1.tot[1:10,]

```

# Model fit results for pooled gender

Note: All factors were adjusted for age and bmi (in the year comparisons).

```{r fit1}

  col.include = c("mod", "Chi-sq", "df", "p-value",
                  "LL", "CFI", "TLI", "BIC", "AIC", "RMSEA",
                  "description")
  mod.2 = mod.1.tot[col.include]
  mod.2$mod = as.integer(mod.1.tot$mod)
  colnames(mod.2)[1] = "Model"
  head(mod.2)

  z = ztable(mod.2, zebra=4, zebra.color="pastelgray")

  # Add row labels
  rowlabs = with(mod.1.tot, {
    unique(paste0("Age: ", Age, ". Groups:", Groups))
  })
  rowlabs

    rowlabs.n = with(mod.1.tot, {
    table(paste0("Age: ", Age, ". Groups:", Groups))
  })

  z = addrgroup(z, rgroup = rowlabs, n.rgroup = rowlabs.n)
  z = update_ztable(z, include.rownames=F, 
                    caption="Model Fit Information")
```

```{r, results='asis'}
z

```

# Test model fit

Use nested chi-square tests to compare models 1 to 3 (some form of variance across groups) to model 4(complete invariance across groups) [@van_de_schoot_checklist_2012]

* **Model 1**. metric invariance: Factor loadings are eq across groups but int differ across groups (metric invariance)
* **Model 2**. Intercept only invariance: factor loadings are different but interceptes are eq across groups
* **Model 3**. Scalar invariance: Same factor loadings and same intercepts for underlying items
* **Model 4**. Full uniqueness invariance: equal loadings and intercept. Also equal variance across all groups

Separate models were run for groupings: RE = racial/ethnic groups, YE = years (2003-2004, 2005-2006, 2007-2008, 2009-2010, 2011-2012)

```{r}

# reshape data with chi-sq values for each model in columns by types of model (age and grouping combos)
head(order)

get.pvals = function(df1) {
  #df1=mod.1.tot
  df1 = df1[complete.cases(df1$df),]
  table(df1$Groups)
  nrow(df1)
  head(df1)
  df1

  chi = dcast(setDT(df1), Age + Groups ~ mod, 
                        value.var = "Chi-sq")

  colnames(chi)[3:6] = c("m1", "m2", "m3", "m4")
  chi

  df = dcast(setDT(df1), Age + Groups ~ mod, 
                        value.var = "df")

  colnames(df)[3:6] = c("m1", "m2", "m3", "m4")
  df

  # get diff in chi sq for model 4 vs 3, 2 and 1
  chi = within(chi, {
    diff.4.vs.1 = m4-m1
    diff.4.vs.2 = m4-m2
    diff.4.vs.3 = m3-m4
  })

  chi

  # get diff in chi sq for model 4 vs 3, 2 and 1
  df = within(df, {
    df.4.vs.1 = m4-m1
    df.4.vs.2 = m4-m2
    df.4.vs.3 = m3-m4
  })

  df = data.frame(df)
  chi = data.frame(chi)

  both.vals = cbind(chi[c("Age", "Groups",
                          "diff.4.vs.1",
                          "diff.4.vs.2",
                          "diff.4.vs.3")],
                    df[c("df.4.vs.1",
                         "df.4.vs.2",
                         "df.4.vs.3")])

  both.vals

  # These models are nested so use difference in chi-sq tests to determine if 
  # any changes are statistically different from the full uniqueness invariance model (model 4)
  both.vals = within(both.vals, {
    pval.m1 = round(pchisq(diff.4.vs.1, df.4.vs.1, lower.tail=F),4)
    pval.m2 = round(pchisq(diff.4.vs.2, df.4.vs.2, lower.tail=F),4)
    pval.m3 = round(pchisq(diff.4.vs.3, df.4.vs.3, lower.tail=F),4)
  })

  return(both.vals)
}

bv = get.pvals(mod.1.tot)

bv.m = get.pvals(mod.1m)
bv.f = get.pvals(mod.1f)

```

```{r}

make.z.tests = function(df.vals, type){
  #df.vals is the data frame
  #type is the type of labeling for the final table

  cgroups = c( "", 
                          "Chi sq diff",
                          "df",
                          "p-value"
                          )

  both.vals.2 = df.vals[, !(colnames(df.vals) %in% c("Age"))]
  colnames(both.vals.2)
  both.vals.2 = both.vals.2[c("Groups", "diff.4.vs.1", "diff.4.vs.2", "diff.4.vs.3",
                              "df.4.vs.1", "df.4.vs.2", "df.4.vs.3",
                              "pval.m1", "pval.m2", "pval.m3")] # reorder order of columns
  colnames(both.vals.2) # check

  colnames(both.vals.2) = c("Groups", rep(c("Model 1 vs 4", "Model 2 vs 4", "Model 3 vs 4"),3))

  n.cgroup = c(1,3,3,3)
  z=ztable(both.vals.2, zebra=0, zebra.color="pastelgray")
  z = addcgroup(z, cgroup=cgroups, n.cgroup=n.cgroup)

  rgroup = c("12 to 19 years", "20 to 50 years", "50 to 80 years")
  if(type==1) {n.rgroup = c(rep(3,3))
                  z = addrgroup(z, rgroup = rgroup, n.rgroup = n.rgroup)}

  if(type==2) {n.rgroup2 = c(rep(2,3))
                z = addrgroup(z, rgroup = rgroup, n.rgroup = n.rgroup2)
  }

  z = update_ztable(z, include.rownames=F,
                    caption="Model fit tests",
                    zebra=2,
                    digits = c(0,0,rep(1,3), rep(0,3), rep(4,3))); #z
}

z.all = make.z.tests(bv, type=1)
z.f = make.z.tests(bv.f, type=2)
z.m = make.z.tests(bv.m, type=2)

```

## Both genders

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
  z.all
```

## Females

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
  z.f
```

## Males

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
  z.m
```

# Model parameters for all models

## Standardized factor loadings by age group, grouping (RE=race/ethnicity and YE=years) and model

```{r}

# YEARS
names(amp.years.params)

# pick off the standardized estimates, stdyx.standardized (3rd element in each list)
# and put in table for all models
names(amp.years.params)  # amp.years.params is a list of lists containing parameters from many different parts of the output for each model, indexed by model.
names(amp.years.params[[1]]$stdyx.standardized) # 7 columns
nrow(amp.years.params[[1]]$stdyx.standardized) # 130 rows
amp.years.params[[1]]$stdyx.standardized

# take out the standardized analysis from each of the model output (3rd element from each list, stdyx.standardized)
#res3 = do.call(rbind.data.frame, sapply(amp.years.params, function(x) x[3]))
#res3$names = rownames(res3)

res3 = rbindlist(sapply(amp.years.params, function(x) x[3]), 
                 use.names=TRUE, fill=TRUE, idcol="names")

head(res3)

# extract out model info (age and group)
#age.group = unlist(strsplit(rownames(res3), "[.]"))
res3$model = substr(res3$names,1,2)
res3$age.group = substr(res3$names,4,9)
res3$grouping = substr(res3$names,11,12)
res3$gender = substr(res3$names, 19, 21)

table(res3$model); table(res3$age.group); table(res3$grouping);
table(res3$gender)
head(res3[res3$Group=="FEMALE",])

# make a new value to put in table: est (se)
res3 = within(res3, {
  est.se = paste0(est, " (", se, ")")
})
head(res3$est.se)

# Look at factor loadings by model

fl = res3[res3$paramHeader %in% c("F1.BY"),]
nrow(fl) # 1065 rows
head(fl)

t1 = data.frame(dcast(setDT(fl[fl$gender %in% c("all", "gen"),]), 
                      age.group + grouping + model + Group ~ param, 
                      value.var = "est.se"))

colnames(t1)[1:4] = c("Age", "Grouping", "Model", "Group value")
head(t1)
table(t1$"Group value")

# females
t1f = data.frame(dcast(setDT(fl[fl$gender=="fem",]), age.group + grouping + model + Group ~ param, 
                      value.var = "est.se"))
colnames(t1f)[1:4] = c("Age", "Grouping", "Model", "Group value")
head(t1f)

# males
t1m = data.frame(dcast(setDT(fl[fl$gender=="mal",]), age.group + grouping + model + Group ~ param, 
                      value.var = "est.se"))
colnames(t1m)[1:4] = c("Age", "Grouping", "Model", "Group value")
head(t1m)

```

```{r}
# function to make standardized estimates in a z table

make.stdest = function(dat){
#dat=t1
  col.include = c("Group value", "HDL2", "LDL2", "TG2")
  col.names = c("Group", "HDL", "LDL", "TG")
  head(t1)
  dat.2 = dat[col.include]
  colnames(dat.2) = col.names
  head(dat.2)

  z = ztable(dat.2, zebra=0, zebra.color="pastelgray")

  # Add row labels
  rowlabs = with(dat, {
    unique(paste0("Age: ", Age, ". Groups:", Grouping, ". Model: ", substr(Model,2,3)))
  })
  rowlabs

  rowlabs.n = with(dat, {
    table(paste0("Age: ", Age, ". Groups:", Grouping, ". Model: ", substr(Model,2,3)))
  })

  z = addrgroup(z, rgroup = rowlabs, n.rgroup = rowlabs.n)
  z = update_ztable(z, include.rownames=F, 
                    caption="Factor Loadings (se)", 
                    zebra=2)
  return(z)
}

z = make.stdest(t1)
zm = make.stdest(t1m)
zf = make.stdest(t1f)

#  z
```

## Pooled gender

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
z
```

## Females

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
zf
```

## Males

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
zm
```

-------

## Standardized residual variance by age group, grouping (RE=race/ethnicity and YE=years) and model

```{r}
resid = res3[res3$paramHeader %in% c("Residual.Variances"),]
nrow(resid)
head(resid)

t2 = data.frame(dcast(setDT(resid[resid$gender %in% c("all", "gen"),]), 
                      age.group + grouping + model + Group ~ param, 
                      value.var = "est.se"))
colnames(t2)[1:4] = c("Age", "Grouping", "Model", "Group value")
t2

# females
t2f = data.frame(dcast(setDT(resid[resid$gender=="fem",]), age.group + grouping + model + Group ~ param, 
                      value.var = "est.se"))
colnames(t2f)[1:4] = c("Age", "Grouping", "Model", "Group value")

# males
t2m = data.frame(dcast(setDT(resid[resid$gender=="mal",]), age.group + grouping + model + Group ~ param, 
                      value.var = "est.se"))
colnames(t2m)[1:4] = c("Age", "Grouping", "Model", "Group value")

```

```{r}

# function to make z table of standardized residuals
makestdres = function(dat) {
  #dat=t2
  head(dat)
  col.include = c("Grouping", "HDL2", "LDL2", "TG2")
  col.names = c("Group", "HDL", "LDL", "TG")
  dat.2 = dat[col.include]
  colnames(dat.2) = col.names
  head(dat.2)
  names(dat)

  z = ztable(dat.2,  zebra=0, zebra.color="pastelgray")

  # Add row labels
  rowlabs = with(dat, {
    unique(paste0("Age: ", Age, ". Groups:", Grouping, ". Model: ", substr(Model,2,3)))
  })
  rowlabs

    rowlabs.n = with(dat, {
    table(paste0("Age: ", Age, ". Groups:", Grouping, ". Model: ", substr(Model,2,3)))
  })

  rgroup = rowlabs

  z = addrgroup(z, rgroup=rowlabs, n.rgroup = rowlabs.n)
  z = update_ztable(z, include.rownames=F, 
                    caption="Standardized residual variance (se)",
                    zebra=2)
}

z = makestdres(t2)
z

zf = makestdres(t2f)
zm = makestdres(t2m)
#  z
```

## Pooled gender

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
z
```

## Females

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
zf
```

## Males

**NOTE**: 
* GE = gender groups
* RE = racial/ethnic groups
* YR = year groups

```{r, results='asis'}
zm
```
----

## Standardized intercepts by age group, grouping (RE=race/ethnicity and YE=years) and model

```{r}
int = res3[res3$paramHeader %in% c("Intercepts"),]
nrow(int)
head(int)

t3 = data.frame(dcast(setDT(int[int$gender %in% c("all", "gen"),]), age.group + grouping + model + Group  ~ param, 
                      value.var = "est.se"))
colnames(t3)[1:4] = c("Age", "Grouping", "Model", "Group value")
head(t3)

t3f = data.frame(dcast(setDT(int[int$gender=="fem",]), age.group + grouping + model + Group  ~ param, 
                      value.var = "est.se"))
colnames(t3f)[1:4] = c("Age", "Grouping", "Model", "Group value")
head(t3f)

t3m = data.frame(dcast(setDT(int[int$gender=="mal",]), age.group + grouping + model + Group  ~ param, 
                      value.var = "est.se"))
colnames(t3m)[1:4] = c("Age", "Grouping", "Model", "Group value")
head(t3m)

```

```{r}
# function to show standardized intercept

makeint = function(dat){

  #dat=t3
  col.include = c("Group value", "HDL2", "LDL2", "TG2")
  col.names = c("Group", "HDL", "LDL", "TG")
  dat.2 = dat[col.include]
  colnames(dat.2) = col.names
  head(dat.2)

  z = ztable(dat.2, zebra=0, zebra.color="pastelgray")

  # Add row labels
  rowlabs = with(dat, {
    unique(paste0("Age: ", Age, ". Groups:", Grouping, ". Model: ", substr(Model,2,3)))
  })
  rowlabs

    rowlabs.n = with(dat, {
    table(paste0("Age: ", Age, ". Groups:", Grouping, ". Model: ", substr(Model,2,3)))
  })
  rowlabs.n

  z = addrgroup(z, rgroup=rowlabs, n.rgroup = rowlabs.n)
  z = update_ztable(z, include.rownames=F, 
                    caption="Standardized intercepts (se)", 
                    zebra=2)
}

z = makeint(t3)
zf = makeint(t3f)
zm = makeint(t3m)
#  z
```

## Pooled gender

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
z
```

## Females

**NOTE**: 
* GE = gender groups
* RE = racial/ethnic groups
* YR = year groups

```{r, results='asis'}
zf
```

## Males

**NOTE**:

* GE = gender groups

* RE = racial/ethnic groups

* YR = year groups

```{r, results='asis'}
zm
```

# References