### 2017-08-21 mplus-results-manuscript.Rmd
initial commit for manuscript results

---
title: "Mplus Results"
author: "Ann Von Holle"
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: lit.bib
output:
  html_document:
        toc: yes
        toc_float: yes
        number_sections: true
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(knitr)
opts_chunk$set(fig.width=10, fig.height=10, 
               fig.keep = 'all',
               fig.align="left", 
               echo=F, results='hide',
               warning=FALSE, message=FALSE, comment=NA)

library("MplusAutomation")
library(reshape2)

library(data.table)
library(plyr)
library(kableExtra)

library(ztable)
options(ztable.type="html")
#options(ztable.type="viewer") # for debugging ztable

```

```{r}

# Read in values
# -------------------

# set directiory to spot where Mplus output for configure testing is stored
setwd("~/../Dropbox/unc.grad.school/my-papers/ms-201608-2/ms-cfa-lipids/mplus/mplus-configural testing")

# Read in information from the .out files
# read in lgmm data info 
output = readModels()
names(output[[1]])
names(output[[1]]$parameters) # get names of different param output.

# get fit summaries from output files
df.sums = lapply(output, function(x) x$summaries)
df.sums[[1]] # check

# combine all model summaries across different groupings into one data frame
combine.sums = do.call("rbind.fill", df.sums)
head(combine.sums)

# make a unique id
combine.sums$id.model = rep(paste0(names(output)),
                              sapply(output, function(x) 
                                nrow(data.frame(x$summaries))))
table(combine.sums$id.model) # check

# Make identifying information for the parameters
handle1 = function(dat) {
dat = within(dat, {
      # set up labels for outcome and sex of child groups (for facetting below)
   groups = ifelse(grepl("groups.age", id.model)==T, "Age",
                 ifelse(grepl(".groups.re.", id.model)==T, "Race/Ethnicity", 
                        ifelse(grepl(".groups.sex.", id.model)==T, "Gender",
                               ifelse(grepl(".groups.year.", id.model)==T, "Time", NA))))
    })
}

combine.sums = handle1(combine.sums)
table(combine.sums$groups) # check

# --------------------------------------------------
# --------------------------------------------------

# get parameters from output files -- for metric models 
# -------------------------------------------------------
# TO DO: need to get standardized estimates
df.params = lapply(output, function(x) data.frame(x$parameters$unstandardized$METRIC.MODEL))
nrow(df.params[[2]])

# Combine all param info across different groupsings into one data frame
combine.params = do.call("rbind.fill", df.params)
nrow(combine.params)

# make a unique id
combine.params$id.model = rep(paste0(names(output)),
                              sapply(output, function(x) 
                                nrow(data.frame(x$parameters$unstandardized$METRIC.MODEL))))
table(combine.params$id.model)

combine.params = handle1(combine.params)
table(combine.params$groups) # check

```

```{r}

# make a table of fit statistics
names(combine.sums)
class(combine.sums)

fits = c("ChiSqM_Value", "ChiSqM_DF",
         "LL", "CFI", "aBIC", "RMSEA_Estimate")

sums.dt = data.table(combine.sums)
wide.fit <- data.table::dcast(sums.dt[sums.dt$Model=="METRIC MODEL", with=T],
                  groups ~ Model , 
                  value.var = fits)

wide.fit

```

```{r, results='markup'}

#dim(wide.fit)
wide.fit2 = wide.fit
colnames(wide.fit2) = c(rep(c(""),7))

kable(wide.fit2, 
      booktabs=T,
      format="html",
      caption='Goodness of fit') %>%
    kable_styling(bootstrap_options = c("striped")) %>%
  add_header_above(c("Groups"=1,
                     "Chi-Square"=1,
                     "DF"=1,
                     "LL"=1,
                     "CFI"=1,
                     "aBIC"=1,
                     "RMSEA"=1)) #%>%
  # group_rows("Weight", 1, 7) %>%
  # group_rows("Height", 8, 14) %>%
  # group_rows("WFL", 15, 21)

```