### 2017-08-23 mplus-results-manuscript.Rmd
updates to cfa ms

---
title: "Mplus Results"
author: "Ann Von Holle"
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: lit.bib
output:
  html_document:
        toc: yes
        toc_float: yes
        number_sections: true
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(knitr)
opts_chunk$set(fig.width=10, fig.height=10, 
               fig.keep = 'all',
               fig.align="left", 
               echo=F, results='hide',
               warning=FALSE, message=FALSE, comment=NA)

library("MplusAutomation")
library(reshape2)

library(data.table)
library(plyr)
library(kableExtra)

library(ztable)
options(ztable.type="html")
#options(ztable.type="viewer") # for debugging ztable

```

```{r}

# Read in values
# -------------------

# set directiory to spot where Mplus output for configural testing is stored
setwd("~/../Dropbox/unc.grad.school/my-papers/ms-201608-2/ms-cfa-lipids/mplus/mplus-configural testing")

# Read in information from the .out files
# read in lgmm data info 
output = readModels()
names(output[[1]])
names(output[[1]]$parameters) # get names of different param output.
```

```{r}
# get fit summaries from output files
df.sums = lapply(output, function(x) x$summaries)
df.sums[[1]] # check

# combine all model summaries across different groupings into one data frame
combine.sums = do.call("rbind.fill", df.sums)
head(combine.sums)

# make a unique id
combine.sums$id.model = rep(paste0(names(output)),
                              sapply(output, function(x) 
                                nrow(data.frame(x$summaries))))
table(combine.sums$id.model) # check

# Make identifying information for the summaries
handle1 = function(dat) {
dat = within(dat, {
      # set up labels for outcome and sex of child groups (for facetting below)
   groups = ifelse(grepl("groups.age", id.model)==T, "Age",
                 ifelse(grepl(".groups.re.", id.model)==T, "Race/Ethnicity", 
                        ifelse(grepl(".groups.sex.", id.model)==T, "Gender",
                               ifelse(grepl(".groups.year.", id.model)==T, "Time", NA))))
    })
}

combine.sums = handle1(combine.sums)
table(combine.sums$groups) # check
```

```{r}

# make a table of fit statistics
names(combine.sums)
class(combine.sums)

fits = c("ChiSqM_Value", "ChiSqM_DF",
         "LL", "CFI", "aBIC", "RMSEA_Estimate")

sums.dt = data.table(combine.sums)
wide.fit <- data.table::dcast(sums.dt,
                  groups + Model ~ . , 
                  value.var = fits)

wide.fit

```

## Model fit statistics

```{r fit, results='markup'}

#dim(wide.fit)
wide.fit2 = wide.fit
wide.fit2$Model = factor(wide.fit2$Model,
                         labels=c("Configural",
                                  "Metric",
                                  "Scalar"))
kable(wide.fit2[,-1], 
      booktabs=T,
      format="html",
      caption='Goodness of fit',
      col.names=c("Model", "ChiSq", "df", "LL", "CFI", "aBIC", "RMSEA")) %>%
    kable_styling(bootstrap_options = c("striped")) %>%
  group_rows("Age", 1, 3) %>%
  group_rows("Gender", 4, 6) %>%
  group_rows("Race/ethnicity", 7, 9) %>%
  group_rows("Time", 10, 12)

```

```{r}

# Read in parameter estimates
# ----------------------------

# set directiory to spot where Mplus output for metric standardized param estimates are
setwd("~/../Dropbox/unc.grad.school/my-papers/ms-201608-2/ms-cfa-lipids/mplus/mplus-configural testing/no-test")

# Read in information from the .out files
# read in lgmm data info 
output.p = readModels()
names(output.p[[1]])
names(output.p[[1]]$parameters) # get names of different param output.
```

```{r}

# --------------------------------------------------
# --------------------------------------------------

# get parameters from output files -- for metric models 
# -------------------------------------------------------

df.params = lapply(output.p, function(x) {
  data.frame(x$parameters$stdyx.standardized)
} )

nrow(df.params[[2]])
df.params[[1]]

# Combine all param info across different groupsings into one data frame
combine.params = do.call("rbind.fill", df.params)
nrow(combine.params)
names(output.p)

# make a unique id
combine.params$id.model = rep(paste0(names(output.p)),
                              sapply(output.p, function(x) 
                                nrow(data.frame(x$parameters))))
table(combine.params$id.model)

# Need to figure out a way to add the groups for each of there unique combos of 
# paramHeader and param.
# order of groups: age (3), re (5), sex (2) and year (5)

# note that the f1.by values are duplicated are the same for factor loadings
# because they are held constant across groups.
# for age: 1 = 12-13 2 = 14-16 3 = 17-19
# for race/ethnicity:  1 = MA 2 = OH 3 = NHW 4 = NHB 5 = OTH
# for sex of child: 1 = male 0 = female
# for year: 2003 = y2003 2005 = y2005 2007 = y2007 2009 = y2009 2011 = y2011

combine.params = handle1(combine.params)
table(combine.params$groups) # check

```

```{r}

# make a table of parameters
head(combine.params)

names(combine.params)
head(combine.params$groups)

wide.params <- data.table::dcast(setDT(combine.params[combine.params$paramHeader=="Intercepts",]),
                  groups + Group  ~ param,
                  value.var = c("est"))

# standardized factor loadings
wide.params.f <- data.table::dcast(setDT(combine.params[combine.params$paramHeader=="F1.BY",]),
                  groups + Group  ~ param,
                  value.var = c("est"))

# groups:
# for age: 1 = 12-13 2 = 14-16 3 = 17-19
# for race/ethnicity:  1 = MA 2 = OH 3 = NHW 4 = NHB 5 = OTH
# for sex of child: 1 = male 0=female
# for year: 003 = y2003 2005 = y2005 2007 = y2007 2009 = y2009 2011 = y2011

wide.params

num.rows.p = table(wide.params$groups)
num.rows.p
num.rows.p.c = cumsum(num.rows.p)

```

## Standardized Intercepts

```{r int, results='markup'}

kable(wide.params[, -1], 
      booktabs=T,
      format="html",
      caption='Standardized Intercepts') %>%
    kable_styling(bootstrap_options = c("striped")) %>%
  group_rows("Age", 1, num.rows.p[1]) %>%
  group_rows("Gender", num.rows.p.c[1]+1, num.rows.p.c[2]) %>%
  group_rows("Race/Ethnicity", num.rows.p.c[2]+1, num.rows.p.c[3]) %>%
  group_rows("Time", num.rows.p.c[3]+1, num.rows.p.c[4])

```

# Standardized Factor Loadings

```{r factorload}

kable(wide.params.f[, -1], 
      booktabs=T,
      format="html",
      caption='Standardized Intercepts') %>%
    kable_styling(bootstrap_options = c("striped")) %>%
  group_rows("Age", 1, num.rows.p[1]) %>%
  group_rows("Gender", num.rows.p.c[1]+1, num.rows.p.c[2]) %>%
  group_rows("Race/Ethnicity", num.rows.p.c[2]+1, num.rows.p.c[3]) %>%
  group_rows("Time", num.rows.p.c[3]+1, num.rows.p.c[4])

```