### 2017-09-07 draft/ms-201705-draft.Rmd
update to draft

---
title: Confirmatory factor analysis (CFA) of multivariate lipid distributions across
  time and race-ethnic groups -- United States, 2003 to 2012
author: "Ann Von Holle, Kari E. North"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    csl: ama.csl
    depth: 3
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
  word_document:
    toc: yes
csl: ama.csl
bibliography: bib1.bib
---
---
title: Confirmatory factor analysis (CFA) of multivariate lipid distributions across time and race-ethnic groups: United States, 2003-2012
author: "Ann Von Holle, Kari E. North"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true # table of content true
    depth: 3  # up to three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
    csl: ama.csl
bibliography: bib1.bib
csl: ama.csl
---

<!-- 
NOTES:

1. This document will be downloaded in text format so any formatting such as bullet points, font changes, bolding, header choices will not be saved. However, please add comments or edits as you see fit. I can clean up later.
2. Any text below preceded by an @ symbol is a placeholder for a citation.
3. To indicate formatting in plain text use the following notation.
Sections in this document start with a # symbol. Section 1 headings start with #, section 2 headings start with ##, etc..
**bold**
*italic*
$latex formula$
4. If you want to see what the text below looks like after formatting, you can cut and paste in https://stackedit.io/editor OR you can go to the web page with a recent version of updates in Word doc format: draft
END OF NOTES
-->

# Abstract

Lipids and lipoproteins found in common profiles such as LDL-C, HDL-C and TG are used to assess dyslipidemia, a well-established risk factor for CVD. These measures are commonly evaluated across socio-demographic groups in a univariate format -- one at a time. Assessing dyslipidemia as a multivariate construct is possible, and perhaps more meaningful, in the context of SEM, but no thorough assessment exists in the literature.

To assess the multivariate construct, we will use confirmatory factor analysis (CFA) to form a latent factor with three commonly assessed lipids and lipoproteins including LDL-C, HDL, TG. In turn, we will assess measurement invariance (MI)  of the construct across demographic characteristics of race/ethnicity and sex in a United States representative sample of children ages 9 to 12 years.

Proper model fit will guide viability of this dyslipidemia construct in analyses. If measurement invariance across racial/ethnic or gender groups exists then there would also be no evidence of differences in the construct itself across these groups and allows for comparisons in means of the construct. Also, similarities across groups also implies the measured lipid indicators have similar pair-wise correlations. Conversely, measurement variance is an indicator that this latent factor is not the same across groups and should be treated as such with stratification.

Results:

We were not able to find evidence of measurement invariance (MI). Tests for MI across racial/ethnic or gender groups are significant indicating that these groups have different multivariate shapes of the three lipid/lipoproteins and the means of these constructs cannot be compared across groups. Within strata of gender or race, the construct has good fit via fit statistics making it a valid construct.

Conclusion:

Analyses demonstrate this construct is a parsimonious way to evaluate dyslipidemia and can provide an alternate approach to treatment of dyslipidemia in structural equation models. However, evaluation of this construct across socio-demographic groups such as race/ethnicity and gender provides evidence of a different multivariate structure requiring stratification.

# Introduction

Dyslipidemia is often characterized through measures of triglycerides (TG), high-density lipoprotein cholesterol (HDL-C), and low-density cholesterol (LDL-C) from a common lipid panel. These measures are then analyzed in research on a univariate basis or with ratios. Another analytic approach can include confirmatory factor analysis (CFA). In practice, this approach would reduce the dimension of three indicators mentioned above to one latent variable capturing the pairwise correlation between measures. In turn, the resulting latent construct can be used either as an exposure or outcome in structural equation models. Although this sort of multivariate analysis can provide a different perspective compared to univariate and traditional approaches, it has not been investigated in the literature as a feasible analytic approach.

CFA is useful for a set of indicators that are interrelated and have already been established in a substantive theoretical sense independent of a statistical rationale. For the three indicators of interest in this particular study: HDL-C, LDL-C, and TG, there exists a strong rationale to treat them together to represent dyslipidemia [@bays_national_2016]. Although there are other indicators of dyslipidemia, these three are the most accessible in common lipid profiles and the most reported in the literature. Any changes in these values represented by a latent factor would be recognizable to lipid researchers and of value in research. What would most likely occur is for TG and LDL-C to be positively correlated and HDL-C negatively correlated with the prior two indicators. A one unit change in the latent factor in this case could represent an increase in TG and LDL-C alongside a decrease in HDL-C given the well established correlations between these measures.

A further motivation to examine dyslipidemia in a dimension reduction approach is to further develop the rationale behind the use of ratios when characterizing lipids. One could consider the correlations between the indicators to be an analog to ratios, with an additional ability to maintain variability for each measure. For example, the use of HDL to TG ratios is increasing in the literature, but analyses using a ratio has limitations [@kuh_correlation_1955; @atchley_ratios_1978; @kronmal_spurious_1993] including failure to adequately adjust for each of the variables forming the ratio when used as a covariate in a linear regression [@kronmal_spurious_1993].  Unlike a ratio, maintaining properties of each lipid measure is possible with a CFA latent construct and perhaps this measure can better inform analyses with this construct as either an outcome or covariate.

Prior to implementation of this approach it is important to investigate if this latent construct represented in the CFA has a good model fit. Poor model fit would indicate that the fitted latent construct does not match the observed data very well and does not serve as a proper latent variable within a numerical dyslipidemia construct. For example, if the indicators are not correlated with one another then this approach is not feasible. Nor is it feasible if the choice of indicators do not fit within a sound theoretical basis. Neither of these cases apply to lipids used in common lipid profiles, and these two requirements make a case for examining dyslipidemia as a latent factor in CFA with the intent of using it in the future either as an exposure or outcome.

Assuming the three measures provide a good source of model fit, a next step would be to evaluate if the form of the latent construct of dyslipidemia stays relatively constant across demographic groups including race/ethnicity, age and gender. Similar form of the construct across different socio-demographic groups allows comparisons of construct means across these groups -- a basic outcome of interest. One common method used to assess this property is measurement invariance (MI) of the latent factor across demographic groups [@van_de_schoot_editorial:_2015-1]. Absent MI then comparing construct means across the demographic groups is more difficult to accomplish. An alternative approach in that resulting scenario would be to stratify analyses for make meaningful inference.

Although the CFA requires additional knowledge of structural equation modeling, the ability to obtain a finer ability to capture correlations between lipid and lipoprotein indicators and condense it into a single latent variable can provide a parsimonious alternative approach to characterize dyslipidemia. Our aim in this study is to assess the fit and test measurement invariance of a latent construct representing dyslipidemia in a U.S.-representative sample. In turn, this assessment and tests can provide evidence of the viability of this construct in future models incorporating dyslipidemia.

# Methods

National Health and Nutrition Examination Survey (NHANES) is a nationwide representative complex survey sample in the United States that assesses population health. Lipid measures are among the data collected in the physical exam [@_nhanes_2017-3]. These analyses include NHANES samples from 2003-2012, spanning ten years. TC and TG were obtained via from venous samples with enzymatic methods [@_nhanes_2017-1; @carroll_trends_2012]. HDL were were measured with a direct immunoassay method [@_nhanes_2017], LDL-C was determined from the Friedewald equation. We included children ages 12 to 19 years in the sample and further subsetted the sample to include the largest racial/ethnic groups in the sample: non-Hispanic White, non-Hispanic Black, and Mexican American. Also, those reporting taking lipid lowering medication were excluded from the sample. After these exclusions there were xx individuals in the complex survey sample.

The primary aim is to create a latent construct of dyslipidemia so we conducted a CFA by specifying one latent variable, dyslipidemia, with three indicators: LDL-C, HDL-C and TG. We specified three factor loadings, one for each indicator, and 3 error variances creating a ‘just identified’ model with zero degrees of freedom [@brown_confirmatory_2015]. All variables required for complex survey sampling were used in the analyses, including the sampling weights, strata and clustering information. We used the Comparative Fit Index, (CFI), adjusted Bayesian Information Criterion (aBIC) and Root Mean Square Error of Approximation (RMSEA) to select the model with best fit. BMI and age are associated with lipid levels and these two factors were allowed to covary with the latent lipid construct to improve model fit.

Measurement invariance was determined by specifying 1) configural, 2) metric and 3) scalar models [@_mplus_2017; @van_de_schoot_checklist_2012]. Socioeconomic groups included age (12-13, 14-15, 16-17, 18-19), gender and racial/ethnic groups. The configural model allows the factor loadings and intercepts to vary across the groups; the metric model holds the factor loading the same across groups; the scalar model holds both the intercept and factor loadings to be the same across groups. To test differences between each of the models we used the chi-square test with a scaling correction factor. If significant, then there was evidence that the more complex model included significant terms that should remain in the model [@brown_confirmatory_2015]. If not significant then we kept the model with fewer parameters.

One aspect that was not possible to incorporate into the model was the method to subset complex survey data given the sampling weights. We did sensitivity tests to determine any bias in ignoring that assumption made on the model parameter estimation and did not detect any difference in effect, considered negligible in substantive meaning.

We used Mplus version [@muthen_mplus_1998] for the CFA analysis and R [@r_core_team_r:_2017] for other data handling tasks.

# Results

For each of the groupings, race/ethnicity and gender, we determined model fit across groups by conducting invariance tests with chi-square tests of three different models against each other. We determined the best fitting model by selecting the most parsimonious model that was no different, i.e. non-significant, from the more complex model.

When comparing gender, racial/ethnic, and age groups, the metric model, allowing intercepts to vary but fixed factor loadings, was not significantly different (Chi-Square=5.6, 2 df, p<0.06; Chi-Square=11.6, 8 df, p>0.16; Chi-Square=8.8, 4 df, p>0.06, respectively) from the configural model, which allowed both loading and intercepts to vary, indicating similarity of factor loadings. On the other hand, the scalar model, fixing both intercepts and factor loadings was different from the configural model (Chi-Square=71.1, 5 df, p<0.0001; Chi-Square=189.2, 20 df, p<0.0001; Chi-Square=86.5, 10 df, p<0.0001, respectively), indicating a difference in factor intercepts. Evidence for different factor intercepts is further supported by a significant test of the scalar versus metric model (Chi-Square=68.8, 3 df, p<0.0001; Chi-Square=198.8, 12 df, p<0.0001, Chi-Square=82.0, 6 df, p<0.0001, respectively). Thus the best fitting model across different gender, racial/ethnic, and age groups is one with similar factor loadings across the groups, but different factor intercepts -- the metric model.

Model fit information (Table 1) was satisfactory for the analyses by gender, racial/ethnic and age groups. First, RMSEA was below 0.05 in all three cases. Second, CFI was above 0.98. Third, aBIC was lowest for the metric model for all three models with the exception of the gender groups, and in this scenario, the difference between the configural and metric model was less than 4 points, considered to be a negligible difference. All of these measures of model fit point towards the metric model having satisfactory model fit and useful for further exploration of the model parameters [@brown_confirmatory_2015].

Parameters of interest for the best fitting model are the intercepts since they are the ones that are allowed to vary across groups in the metric specification of the model. In table 2, the intercepts for the confirmatory factor representing dyslipidemia are described by groups within each separate model. Within the model that separates by age, the three older age groups appear to have shift in the center of the factor in the negative direction for HDL, with standardized intercept resting at -0.163 for the 14 to 16 and 17 to 19 year age groups. For the middle age group, ages 14 to 16, LDL and TG are shifted further to the negative direction and the 17 to 19 year old group are shifted more in the positive direction when compared to the youngest age group 12 to 13 years of age.

When examining the model with gender groups, the males appear to have a shift in the negative direction for HDL and towards a higher value of TG compared to females. Finally, the racial/ethnic groups also have differences in their intercepts. Using the Mexican American (MA) group as a referent, the non-Hispanic Blacks (NHB) have a center for the three-dimensional factor that is shifted more towards a higher HDL and LDL along with a lower TG. The non-Hispanic White group does not appear to have as large a difference as the NHB group. The other Hispanic (OH) group appears to have a larger negative shift in LDL and TG as well as the “Other” racial/ethnic group displaying characteristics similar to the NHB but not as extreme.

Finally, factor loadings are not allowed to vary across groups in the metric model. Factor loadings indicate the degree to which the measured correlation between the indicators are correlated with the dyslipidemia factor concurrently separating out the measurement error -- an advantage of CFA. Examining these association shows that TG has the strongest correlation with the factor, at a correlation exceeding 0.71. HDL has the second strongest association -- this time a negative one -- around -0.52. Finally, LDL is the least correlated with the factor with a correlation at 0.39 for this sample. Any standardized factor loading below 0.3 indicates need to remove that indicator, and none of the lipid indicators fit that description. These values culminate in an interpretation of a one unit increase in the factor equating with an increase in adverse levels of lipids: declining HDL and increasing TG and LDL.

# Discussion

The purpose of this paper was to examine the use of CFA as a measure of dyslipidemia. First, we evaluated measures of model fit of a CFA with three common lipid measures serving as observed indicators in the model: HDL-C, LDL-C and TG. Nested within this aim is the concept that this CFA dyslipidemia construct may not be similar across sociodemographic groups such as race/ethnicity and age -- a common area of comparisons for lipid measures. Given known differences in the factor across these demographic categories, we tested this latter hypothesis through the measurement invariance (MI) concept. MI did not hold for different racial/ethnic nor to age groups indicating that the factor means can not be meaningfully compared across these groups, which has implications for future use.

## Measurement Invariance

In evaluating the three different types of models across subgroups for each of the following domains: age, race/ethnicity and sex of child, the metric model was the best of the three for all domains. The configural model, in which all factor loadings and means of the dyslipidemia factor were allowed to vary across groups, was not statistically significantly different from the metric model. Given this result, the metric model offers a parsimonious fit by fixing the factor loadings to be the same across groups. Furthermore, the metric model was different from the scalar model, adding more evidence to support the choice of the metric model.

Similar factor loadings across the subgroups implies correlations between the three lipid indicators are the same across the subgroups within each domain. This finding may have relevance for lipid ratio measures such as HDL to TG, a commonly used metric. For example, a higher factor score for a construct with positive HDL-C factor loadings combined with negative LDL-C and TG factor loadings would mean on average a simultaneous increase of HDL and decrease in LDL-C and TG, equivalent to a higher HDL-C to LDL-C ratio. This interpretation implies higher ratios are analogous to higher factor score as well as lower ratios having lower factor scores.

Although the factor loadings for the construct are similar, this finding does not imply metric invariance holds for these subgroups. In fact, the measurement invariance testing does not support comparing factor means across subgroups within each domain. Metric invariance implies the centers of this multidimensional construct are different across the subgroups. For example, when the factor is equal to zero for all subgroups, the estimated indicator (HDL-C, LDL-C and TG) are all different depending on the subgroup you choose. This result has implications regarding use of mean factor scores across subgroups. Some researchers tends to obtain factor scores and use them as observed entities, then comparing them across subgroups. These analyses demonstrate the perils in this approach with scientifically meaningless results given the evidence of scale differences. The existence of measurement non-invariance supports different dyslipidemia constructs for each of the selected subgroups for race/ethnicity, age and gender. As such, comparisons in factor score means have different meanings and are not substantively meaningful. Although this finding remains restricted to this example, it pertains to a nationwide representative sample of children in the United States and provides an important reminder when conducting analyses in other samples.

## Fit

Finally, having chosen the metric model for all of the three domains, we determined that model fit diagnostics were within normal bounds of adequate to good fit and as such motivates further use of this construct. Use of this construct would be an interesting supplement to univariate outcomes with a different and similarly meaningful interpretation.

Measuring dyslipidemia as one joint distribution that incorporates the correlation between each of three commonly measured lipids is worthwhile either as an outcome or exposure in research. Lipids do not function independently of one another and a measure addressing their covariance is an asset. CFA leverages this characteristic of these measures to form a measure with different meaning from univariate indicators. Combining the observed correlations between the three common lipids alongside their common presentation in the literature and theoretical basis for dyslipidemia further justifies the use of CFA.

This alternate construct can be an adjunct measure to univariate measures of lipids, which provides a means to assess the degree to which one measure increases as another measure either increase or decreases on a population level. Aside from the important caveat of measurement non-invariance, this work demonstrates the degree to which HDL-C, LDL-C and TG are correlated with one another on a population level. In turn this correlation can provide meaningful associations with other factors in downstream analyses.

# References